---
title: "Management Strategy Evaluation of Biomass Dynamic Management Procedures"
author: "Laurence Kell"
date: "August 13th, 2016"
output:
    tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
bibliography: /home/laurie/bib/refs.bib
link-citations: yes
vignette: >
  %\VignetteIndexEntry{mpb}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r knitr_init, echo=FALSE, results="hide"}
library(knitr)
## Global options
opts_chunk$set(echo    =TRUE,
               eval    =TRUE,
               cache   =!FALSE,
               cache.path="cache/",
               prompt  =FALSE,
               comment =NA,
               message =FALSE,
               tidy    =FALSE,
               warnings=FALSE,
               fig.height=4.5,
               fig.width =4.5,
               fig.path  ="tex/")
```

```{r,echo=FALSE}
sink(NULL)
warn=options()$warn
options(warn=-1)
library(ggplotFL)
library(plyr)
library(reshape)
library(mpb)

theme_set(theme_bw())
options(digits=3)
options(warn=warn)
sink()
```

# Introduction

\newthought{Management Strategy Evaluation (MSE)} requires six steps @punt2007developing; namely i) identification of management objectives; ii) selection of hypotheses for the OM; iii) conditioning the OM based on data and knowledge, and possible weighting and rejection of hypotheses; iv) identifying candidate management strategies; v) running the Management Procedure (MP) as a feedback control in order to simulate the long-term impact of management; and then vi) identifying the MPs that robustly meet management objectives. 

In this document we provide an example of conducting and MSE using the `mpb` package.


```{r}
library(mpb)
library(FLash)
library(FLBRP)
```

# Management objectives 
# Operating Model 
# Management strategies 
# Running 
the Management Procedure (MP) as a feedback control in order to simulate the long-term impact of management;  
# Identifying the MPs that robustly meet management objectives. 


```{r}
# load(om)
# load(eql)
load("/home/laurie/Desktop/flr/mpb/data/om.RData")
load("/home/laurie/Desktop/flr/mpb/data/eql.RData")
```

```{r}
#' om=fwdWindow(om,eql,end=2030)
#' om=propagate(om,100)
```

```{r}
#' srDev=FLQuant(0,dimnames=list(year=2000:2030))
#' srDev=rlnorm(100,srDev,0.3)
#' 
#' om=fwd(om,catch=catch(om)[,ac(2000:2011)],sr=eql,sr.residuals=srDev)
#' om=fwdWindow(om,eql,end=2040)
```

```{r}
#' 
#' uDev=FLQuant(0,dimnames=list(year=ac(dims(eql)["minyear"]):ac(dims(eql)["maxyear"])))
#' uDev=rlnorm(100,uDev,0.3)
#' 
#' library(popbio)
#' mp=FLBRP2biodyn(  eql,"biodyn")
#' 
#' modelParams=mpb:::modelParams
#' setParams(mp)=ssb.obs(eql)%*%uDev
#' setControl(mp)=params(mp)
#' 
```

# References