---
title: "Extended Survivors Analysis"
author: "Laurence Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
bibliography: /home/laurie/bib/refs.bib
tags: FLPKG FLR
license: Creative Commons Attribution-ShareAlike 4.0 International
---

```{r knitr_init, echo=FALSE, results="hide"}
library(knitr)
## Global options
opts_chunk$set(echo    =!TRUE,
               eval    =!TRUE,
               cache   =!FALSE,
               cache.path="cache/",
               prompt  =FALSE,
               comment =NA,
               message =FALSE,
               tidy    =FALSE,
               warnings=FALSE,
               fig.height=4.5,
               fig.width =4.5,
               fig.path  ="tex/")
```

```{r, pkgs, echo=FALSE, message=FALSE}
sink(NULL)
options(warn=-1)
library(FLCore)
library(FLXSA)
library(ggplotFL)

theme_set(theme_bw())
options(digits=3)
options(warn=warn)
sink()
```

# Introduction

@shepherd1999extended
@kell2007flr
@kell2005flat
@kell2005round

# Background / Methods

This summation expression $\sum_{i=1}^n X_i$ appears inline.

This summation expression is in display form.

$$\sum_{i=1}^n X_i$$

# Using the FLPKG package

```{r class1,eval=TRUE,echo=TRUE}
data(ple4)
data(ple4.indices)
```

```{r class2,echo=TRUE,eval=FALSE}
is(ple4)
getSlots(is(ple4))
```

```{r class3,echo=TRUE,eval=FALSE}
is(ple4.indices)
getSlots(is(ple4.indices[[1]]))
```

## Plotting

\newthought{Plots} are important for examining objects, exploring data, summarising results, checking outputs, and diagnosing problems. FLXSA uses the `ggplot2` and `ggplotFL` packages.

```{r plot,echo=TRUE,eval=FALSE}
library(ggplotFL)
```

```{r plot2, eval=TRUE, echo=TRUE, fig.margin=TRUE, fig.width=4, fig.height=5, fig.cap="Time series"}
library(FLCore)
data(ple4)
plot(ple4)
```

```{r plot3, eval=TRUE, echo=TRUE, fig.margin=TRUE, fig.width=4, fig.height=4, fig.cap="CPUE Time series"}
library(ggplotFL)
data(ple4.indices)
plot(ple4.indices[["SNS"]]@index)+
       geom_point()+
       geom_smooth(se=FALSE)
```

# `FLXSA` Class

## Fitting

```{r fit, echo=TRUE}
ctl     =FLXSA.control()
ple4.xsa=FLXSA(ple4,ple4.indices)
ple4    =ple4+ple4.xsa
```


```{r U16}
plot(ple4)+
  theme_bw()+
  theme(legend.position="none")
```
**Figure `r iFig=1;iFig`**. XSA fits.

```{r diags}
plot(ple4.xsa)
```
**Figure `r iFig=iFig+1;iFig`**. XSA plots, weights for terminal year Ns for each CPUE observation and shrinkage

## Diagnostics

```{r diag, echo=TRUE}
library(diags)

dgs=diags(ple4.xsa)
```


```{r diags2}
source('~/Desktop/flr/FLXSA/R/FLXSA-diags.R')

dgs=diags(ple4.xsa)

ggplot(dgs,aes(hat,residual))+
  geom_point()+
  geom_hline(aes(yintercept=0))+
  facet_grid(name~age,scale="free_x")
```
**Figure `r iFig=iFig+1;iFig`**. XSA diagnostics, residuals against fitted value.

```{r diags3}
ggplot(dgs,aes(year,residual))+
  geom_point()+
  geom_errorbar(aes(year,ymin=0,ymax=residual))+
  geom_smooth(span=1,se=F)+
  facet_grid(name~age)
```
**Figure `r iFig=iFig+1;iFig`**. XSA diagnostics, residuals against year.

```{r diags4,fig.height=2}
ggplot(subset(dgs,age==4),aes(hat,obs))+
  geom_abline(aes(intercept=0,slope=1))+
  geom_point()+
  geom_smooth(method="lm",se=FALSE)+
  facet_wrap(~name,scale="free",ncol=3)+
  theme_bw()
```
**Figure `r iFig=iFig+1;iFig`**. XSA diagnostics, Calibration regression plots for age 4.


```{r diags5,fig.height=6}
ggplot(dgs,aes(residual,residualLag))+
  geom_hline(aes(yintercept=0))+
  geom_point()+
  geom_smooth(method="lm",se=F)+
  facet_grid(age~name,scale="free")+
  theme_bw()
```
**Figure `r iFig=iFig+1;iFig`**. XSA diagnostics, AR plots of lagged residuals

```{r diags6,fig.height=4}
ggplot(dgs)                                           +
  geom_point( aes(qqx,qqy))                            +
  stat_smooth(aes(qqx,qqHat),method="lm",se=T,fill="blue", alpha=0.1)         +
  facet_grid(age~name)                                    +
  theme(legend.position="bottom")  
```
**Figure `r iFig=iFig+1;iFig`**. XSA diagnostics, QQ plots to check for normality


# Uncertainty

```{r uncertainty, echo=TRUE, fig.margin=TRUE}
xsa =FLXSA(ple4,ple4.indices)
ple4=FLXSA:::rand(100,ple4,xsa)
```

## Parameters
## Retrospectives

```{r Retro}
plot(FLStocks(mlply(data.frame(year=2005:2008), function(year,stk,u,ctl){
  stk=window(stk,end=year)  
  stk+FLXSA(stk,u,ctl)
  },stk=ple4,u=ple4.indices,ctl=ctl)))+
  theme_bw()+theme(legend.position="bottom")+
  scale_colour_discrete(labels=2005:2008,guide=guide_legend(title="Retrospective\nYear"))
```
**Figure `r iFig=iFig+1;iFig`**. Retrospective XSA time series estimates.

## Crossvalidation

@kell2016evaluation

# Reference Points
```{r refs, echo=TRUE}
library(FLBRP)
```

```{r sr}
ple4.sr=as.FLSR(ple4,model="bevholt")
ple4.sr=fmle(ple4.sr,control=list(trace=0))
plot(ple4.sr)+
  theme_bw()
```
**Figure `r iFig=iFig+1;iFig`**. Stock Recruitment Relationship

# Projection

```{r fdwd, fig.margin=TRUE,fig.width=4, fig.height=6,fig.cap="Projection"}
``` 

```{r}
library(FLBRP)

brp=FLBRP(ple4,sr=ple4.sr)

plot(brp)+
  theme_bw()+
  theme(legend.position="bottom")
```
**Figure `r iFig=iFig+1;iFig`**. Equilibrium Analysis with reference points.

```{r}
plot(brp,obs=TRUE,refpts=FALSE)+
  theme_bw()+
  theme(legend.position="bottom")
```
**Figure `r iFig=iFig+1;iFig`**. Equilibrium Analysis with observations.


```{r}
library(kobe)

kb=kobe(brp,ple4)
```

```{r}
kobePhase(kb,xlim=c(0,2))
  geom_path(aes(stock,harvest,col=year))+
  theme(legend.position="none")
```
**Figure `r iFig=iFig+1;iFig`**. Kobe Phase Plot.

# Harvest Control Rules
```{r hcr,fig.margin=TRUE,fig.width=6,fig.height=6}
library(FLash)
```

```{r kobe,fig.margin=TRUE,fig.width=6,fig.height=6}
```

# Advice
```{r advice, echo=TRUE}
library(kobe)
```

# MSE

```{r mse,eval=FALSE}
library(mpb)
```


Tables
=====================================

Table 1a. XSA Control options from continuity run.
```{r}
ctl

```


Table 2a. XSA diagnostics from continuity run.
```{r}
diagnostics(ple4.xsa)
```

# More information

* You can submit bug reports, questions or suggestions on `FLPKG` at the `FLPKG` issue page ^[<https://github.com/flr/FLPKG/issues>], or on the *FLR* mailing list.
* Or send a pull request to <https://github.com/flr/FLPKG/>
* For more information on the FLR Project for Quantitative Fisheries Science in R, visit the FLR webpage ^[<http://flr-project.org>].
* The latest version of `FLPKG` can always be installed using the `devtools` package, by calling
```{r, devtools, echo=TRUE, eval=FALSE}
	library(devtools)
	install_github('flr/FLPKG')
```


## Software Versions

* `r version$version.string`
* FLCore: `r packageVersion('FLCore')`
* FLPKG: `r # packageVersion('FLPKG')`
* **Compiled**: `r date()`
* **Git Hash**: `r system("git log --pretty=format:'%h' -n 1", intern=TRUE)`

## Author information

**Name SURNAME**. Affiliation. Address, Locality (RE) Country.

# References
